\begin{problem}[1]
  Let $V$ be a finite-dimensional inner product space over $\C$, and let $T$ be
  a linear transformation on $V$. Prove that $T$ is self-adjoint if and only if
  $\langle \x, T\x \rangle$ is real a real number for all $\x \in V$.
\end{problem}

\begin{proof}[Solution]
\end{proof}

\begin{problem}[2]
  Let $A \in \C^{n \times n}$. Prove that $A$ is normal if and only if $A$ can
  be written in the form of $A = A_1 + iA_2$ where $A_1$ and $A_2$ are Hermitian
  and $A_1A_2 = A_2A_1$.
\end{problem}

\begin{proof}[Solution]
\end{proof}

\begin{problem}[3]
  Prove that a normal and nilpotent linear transformation is the zero linear
  transformation.

  \noindent(Note: A linear transformation $T$ is nilpotent if there exists a
  positive integer $r$ such that $T^r = 0$.))
\end{problem}

\begin{proof}[Solution]
\end{proof}

\begin{problem}[4]
  Let $A \in \C^{n \times n}$ be a Hermitian matrix. Prove that $A$ is positive
  definite if and only if all the eigenvalues of $A$ are positive.
\end{problem}

\begin{proof}[Solution]
\end{proof}

\begin{problem}[5]
  Let $A \in \C^{n \times n}$ be a normal matrix. Prove that $A$ is unitary if
  and only if every eigenvalue of $A$ has magnitude (or absolute value) $1$.
\end{problem}

\begin{proof}[Solution]
\end{proof}

\begin{problem}[6]
  Let $A \in \C^{n \times n}$ be a Hermitian matrix. Suppose $A$ is both
  positive definite and unitary. Prove that $A = I$. \textit{Hint: You may use
  conclusions from the above two problems.}
\end{problem}

\begin{proof}[Solution]
\end{proof}

\begin{problem}[7]
  Consider $A = \begin{pmatrix}
    -1 & 0 & 1 \\
    1 & -1 & 0 \\
  \end{pmatrix}$
  \begin{enumerate}
    \item Find the singular value decomposition of $A$. (Show your computation
      work, do not use technology.)

    \item Find the generalized inverse $A^+$.
  \end{enumerate}
\end{problem}

\begin{proof}[Solution to (i)]
\end{proof}

\begin{proof}[Solution to (ii)]
\end{proof}

\begin{problem}[8]
  Let $A \in \C^{n \times n}$ be a Hermitian and indefinite (i.e. not positive
  definite or positive semi-definite). Suppose $A = PDP^*$ for some unitary
  matrix $P$ and some diagonal matrix $D \in \R^{n \times n}$. Find the singular
  value decomposition of $A = V\Sigma U^*$ by constructing $V$, $\Sigma$, and
  $U$ using some possible variations of the columns or entries of $P$ and $D$.
\end{problem}

\begin{proof}[Solution]
\end{proof}

\begin{problem}[9]
  Let $A$ be an $m \times n$ matrix, and let $P$ and $Q$ be $m \times m$ and $n
  \times n$ unitary matrices. Show that $A$ and $PAQ$ have the same singular
  values.
\end{problem}

\begin{proof}[Solution]
\end{proof}
