\begin{problem}
  Consider the following regions of $\R^3$ described by an inequality. Describe
  each region in words and sketch it to the best of your ability. Note:
  Inequalities define a solid region in space rather than a surface.
  \begin{enumerate}
    \item $4 < x^2 + y^2 + z^2$.

    \item $x^2 + z^2 \le 9$.
  \end{enumerate}
\end{problem}

\begin{probsolution}
  \begin{enumerate}
    \item The inequality represents the points $(x, y, z)$ whose distance from
      the origin is more than $2$. This creates a sphere of radius $2$ centered
      at the origin (see \cref{fig:1_a}). All the points are outside the sphere.

    \item Since $y = 0$ in this inequality, that means the inequality is being
      drawn on the $xz$-plane. This is just a circle of radius $3$ centered at
      the origin (see \cref{fig:1_b})
  \end{enumerate}
\end{probsolution}

\newpage

\begin{problem}
  Consider the vectors $\a = \langle -2, 5, 4 \rangle$ and $\b = \langle 4, 8, 1
  \rangle$.
  \begin{enumerate}
    \item Find $2\a - 3\b$.

    \item Find the length of the vectors $\a$ and $\b$.

    \item Find the angle between the vectors $\a$ and $\b$.

    \item Find a unit vector in the direction of $\a$.

    \item Find $\comp_{\a}(\b)$ and $\proj_{\a}(\b)$.
  \end{enumerate}
\end{problem}

\begin{probsolution}
  \begin{enumerate}
    \item $\displaystyle2\langle -2, 5, 4 \rangle - 3\langle 4, 8, 1 \rangle =
      \langle -4, 10, 8 \rangle - \langle 12, 24, 3 \rangle = \langle -16, -14,
      5 \rangle$.

    \item $\displaystyle\lvert a \rvert = \sqrt{(-2)^2 + 5^2 + 4^2} =
      \sqrt{45}$.

      $\displaystyle\lvert b \rvert = \sqrt{4^2 + 8^2 + 1^2} = \sqrt{81}$.

    \item $\displaystyle\frac{\a \cdot \b}{\lvert \a \rvert \cdot \lvert \b
      \rvert} = \frac{36}{\sqrt{45} \cdot \sqrt{81}} = \frac{4\sqrt{5}}{15}$.
      This gives us $\displaystyle\arccos\left(\frac{4 \sqrt{5}}{15}\right)
      \approx \SI{54}{\degree}$.

    \item The unit vector for $\a$ is $\displaystyle\frac{\a}{\lvert \a \rvert}
      = \frac{\langle -2, 5, 4 \rangle}{\sqrt{45}} = \left\langle
      -\frac{2}{\sqrt{45}}, \frac{5}{\sqrt{45}}, \frac{4}{\sqrt{45}}
      \right\rangle$.

    \item $\displaystyle\comp_{\a}(\b) = \frac{\a \cdot \b}{\lvert \a \rvert} =
      \frac{36}{\sqrt{45}}$.

      $\displaystyle\proj_{\a}(\b) = \frac{\a \cdot \b}{\lvert \a \rvert^2} \a =
      \frac{36}{45} \langle -2, 5, 4 \rangle = \left\langle -\frac{8}{5}, 4,
      \frac{16}{5} \right\rangle$.
  \end{enumerate}
\end{probsolution}

\newpage

\begin{problem}
  Find the values of $x$ such that the vectors $\langle 6, 3x, x \rangle$ and
  $\langle -3, 1, x \rangle$ are orthogonal.
\end{problem}

\begin{probsolution}
\end{probsolution}

\newpage

\begin{problem}
  If $\r = \langle x, y, z \rangle$ is an arbitrary position vector and $\u =
  \langle u_1, u_2, u_3 \rangle$ and $\v = \langle v_1, v_2, v_3 \rangle$ are
  constant position vectors, show that $(\r - \u) \cdot (\r - \v) = 0$ defines a
  sphere. Find the center and radius of the sphere.
\end{problem}

\begin{probsolution}
\end{probsolution}

\newpage

\begin{problem}
  Show that if $\u + \v$ and $\u - \v$ are orthogonal, then $\u$ and $\v$ have
  the same length.
\end{problem}

\begin{probsolution}
\end{probsolution}
