\lecture{16}{Nov 6 2024 Wed (13:02:06)}{Conversion Tests and Double Sums}

\subsection{Conversion Tests}

We finalize this chapter with a collection of conversion tests for series (total of $11$). These tests are essential tools in analysis, allowing us to determine whether a given series converges or diverges without having to compute the sum explicitly.

\begin{lemma}
  If the series $\sum_{k=1}^\infty a_k$ converges, then $a_k \to 0$.
\end{lemma}

\begin{proof}
  Let $S_n = \sum_{k=1}^n a_k$. Then, $a_n = S_n - S_{n-1}$. Since the series converges, the sequence of partial sums is Cauchy. Thus, given $\epsilon > 0$, there exists an $N \in \N$ such that whenever $n > m > N$, it follows that
  \[%
    \lvert S_n - S_m \rvert = \lvert a_{m+1} + a_{m+2} + \cdots + a_n \rvert < \epsilon
  .\]%
  In particular, if we take $m = n - 1$, we have
  \[%
    \lvert a_n \rvert = \lvert S_n - S_{n-1} \rvert < \epsilon
  .\]%
  Thus, $a_n \to 0$.
\end{proof}

\begin{note}
  The converse isn't true. Take, for example, the harmonic series $\sum_{k=1}^\infty \frac{1}{k}$. This series diverges, but $\frac{1}{k} \to 0$.
\end{note}

\begin{lemma}[Absolute Convergence Test]
  If $\sum_{k=1}^\infty \lvert a_k \rvert$ converges, then $\sum_{k=1}^\infty a_k$ converges.
\end{lemma}

\begin{proof}
  Assume $\sum_{k=1}^\infty \lvert a_k \rvert$ converges. Let $\epsilon > 0$. Then, there exists an $N \in \N$ such that for every $n > m > N$, we have
  \[%
    \lvert a_{m+1} \rvert + \lvert a_{m+2} \rvert + \cdots + \lvert a_n \rvert < \epsilon
  .\]%
  Notice that
  \[%
    \lvert a_{m+1} + a_{m+2} + \cdots + a_n \rvert \leq \lvert a_{m+1} \rvert + \lvert a_{m+2} \rvert + \cdots + \lvert a_n \rvert < \epsilon
  .\]%
  Thus, by the Cauchy Criterion for Series, $\sum_{k=1}^\infty a_k$ converges.
\end{proof}


\begin{lemma}[Comparison Test]
  Assume $(a_k)$ and $(b_k)$ are sequences satisfying $(\forall k \in \N)[0 \le a_k \le b_k]$
  \begin{enumerate}
    \item If $\sum_{k=1}^\infty b_k$ converges, then $\sum_{k=1}^\infty a_k$ converges.

    \item If $\sum_{k=1}^\infty a_k$ diverges, then $\sum_{k=1}^\infty b_k$ diverges.
  \end{enumerate}
\end{lemma}

\begin{proof}\leavevmode
  Both statements follow directly from the Cauchy Criterion for Series.
  \[%
    \lvert a_{m+1} + a_{m+2} + \cdots + a_n \rvert \le \lvert b_{m+1} + b_{m+2} + \cdots + b_n \rvert
  .\]%
  If the right-hand side is less than $\epsilon$, then the left-hand side is less than $\epsilon$. Thus, the series converges. The second statement follows from the contrapositive of the first statement.
\end{proof}

\begin{lemma}[Limit Comparison Test]
  Assume $(a_k)$ and $(b_k)$ are sequences satisfying $(\forall k \in \N)[0 < a_k \le b_k]$. If
  \[%
    \lim_{k \to \infty} \frac{a_k}{b_k} = c
  ,\]%
  where $c$ is a positive real number, then $\sum_{k=1}^\infty a_k$ converges if and only if $\sum_{k=1}^\infty b_k$ converges.
\end{lemma}

\begin{proof}
  Let $c > 0$ such that $\lim_{k \to \infty} \frac{a_k}{b_k} = c$. This means for any $\epsilon > 0$, there exists an $N \in \N$ such that for all $k \geq N$, we have
  \[%
    c - \epsilon < \frac{a_k}{b_k} < c + \epsilon
  .\]%
  Multiplying through by $b_k$ (which is positive), we get
  \[%
    (c - \epsilon)b_k < a_k < (c + \epsilon)b_k
  .\]%
  Choose $\epsilon$ such that $0 < \epsilon < c$, so $c - \epsilon > 0$. For $k \geq N$, we now have
  \[%
    0 < (c - \epsilon) b_k \leq a_k \leq (c + \epsilon) b_k
  .\]%

  \begin{enumerate}
    \item Convergence of $\sum_{k=1}^\infty b_k$:

      If $\sum_{k=1}^\infty b_k$ converges, then $(c - \epsilon)\sum_{k=1}^\infty b_k$ converges because $c - \epsilon > 0$ is a constant multiple. By the Comparison Test applied to $(c - \epsilon)b_k \leq a_k$, it follows that $\sum_{k=1}^\infty a_k$ also converges.

    \item Divergence of $\sum_{k=1}^\infty b_k$:

      If $\sum_{k=1}^\infty b_k$ diverges, then $(c + \epsilon)\sum_{k=1}^\infty b_k$ also diverges because $c + \epsilon > 0$ is a constant multiple. By the Comparison Test applied to $a_k \geq (c - \epsilon)b_k$, it follows that $\sum_{k=1}^\infty a_k$ also diverges.
  \end{enumerate}

  Hence, $\sum_{k=1}^\infty a_k$ converges if and only if $\sum_{k=1}^\infty b_k$ converges.
\end{proof}

\begin{lemma}[Ratio Test]
  Let $(a_n)$ be a sequence. If
  \[%
    \lim_{n \to \infty} \left\lvert \frac{a_{n+1}}{a_n} \right\rvert = L
  ,\]%
  then
  \begin{enumerate}
    \item If $L < 1$, then $\sum_{n=1}^\infty a_n$ converges absolutely.

    \item If $L > 1$, then $\sum_{n=1}^\infty a_n$ diverges.

    \item If $L = 1$, then the test is inconclusive.
  \end{enumerate}
\end{lemma}

\begin{proof}\leavevmode
  Let $L = \lim_{n \to \infty} \left\lvert \frac{a_{n+1}}{a_n} \right\rvert$.
  \begin{enumerate}
    \item $L < 1$:

      Choose $\epsilon > 0$ such that $L + \epsilon < 1$. By the definition of the limit,
      \[%
        (\exists N \in \N)(\forall n \ge N)\left[\left\lvert \frac{a_{n+1}}{a_n} \right\rvert < L + \epsilon\right]
      .\]%
      This implies that for $n \geq N$
      \[%
        \lvert a_{n+1} \rvert < (L + \epsilon) \lvert a_n \rvert
      .\]%
      Iterating this inequality, we get
      \[%
        \lvert a_{n+1} \rvert < (L + \epsilon)^{n-N+1} \lvert a_N \rvert
      .\]%
      Since $L + \epsilon < 1$, the geometric series $\sum_{n=N}^\infty (L + \epsilon)^{n-N+1}$ converges. Therefore, by the Comparison Test, $\sum_{n=N}^\infty \lvert a_n \rvert$ converges, which implies that $\sum_{n=1}^\infty \lvert a_n \rvert$ converges. Thus, $\sum_{n=1}^\infty a_n$ converges absolutely.

    \item $L > 1$:

      Choose $\epsilon > 0$ such that $L - \epsilon > 1$. By the definition of the limit,
      \[%
        (\exists N \in \N)(\forall n \ge N)\left[\left\lvert \frac{a_{n+1}}{a_n} \right\rvert > L - \epsilon\right]
      .\]%
      This implies that for $n \geq N$
      \[%
        \lvert a_{n+1} \rvert > (L - \epsilon) \lvert a_n \rvert
      .\]%
      Iterating this inequality, we get
      \[%
        \lvert a_{n+1} \rvert > (L - \epsilon)^{n-N+1} \lvert a_N \rvert
      .\]%
      Since $L - \epsilon > 1$, the terms $\lvert a_n \rvert$ do not approach zero, and hence $\sum_{n=1}^\infty \lvert a_n \rvert$ diverges. Thus, $\sum_{n=1}^\infty a_n$ diverges.

    \item $L = 1$:

      The test provides no information about the convergence or divergence of the series. For example, the series $\sum_{n=1}^\infty \frac{1}{n}$ diverges, but $\sum_{n=1}^\infty \frac{1}{n^2}$ converges, and both satisfy $\lim_{n \to \infty} \left\lvert \frac{a_{n+1}}{a_n} \right\rvert = 1$. Therefore, the test is inconclusive in this case. \qedhere
  \end{enumerate}
\end{proof}

\begin{lemma}[Root Test]
  Let $(a_n)$ be a sequence. If
  \[%
    \limsup_{n \to \infty} \sqrt[n]{\lvert a_n \rvert} = L
  ,\]%
  then
  \begin{enumerate}
    \item If $L < 1$, then $\sum_{n=1}^\infty a_n$ converges absolutely.

    \item If $L > 1$, then $\sum_{n=1}^\infty a_n$ diverges.

    \item If $L = 1$, then the test is inconclusive.
  \end{enumerate}
\end{lemma}

\begin{proof}
  Let $L = \limsup_{n \to \infty} \sqrt[n]{\lvert a_n \rvert}$.

  \begin{enumerate}
    \item $L < 1$:

      Choose $\epsilon > 0$ such that $L + \epsilon < 1$. By the definition of the $\limsup$, there exists $N \in \N$ such that for all $n \geq N$
      \[%
        \sqrt[n]{\lvert a_n \rvert} < L + \epsilon
      .\]%
      Raising both sides to the $n$th power, we have
      \[%
        (\forall n \geq N)\left[\lvert a_n \rvert < (L + \epsilon)^n\right]
      .\]%
      Since $L + \epsilon < 1$, the series $\sum_{n=N}^\infty (L + \epsilon)^n$ converges because it is a geometric series with ratio less than $1$. By the Comparison Test, $\sum_{n=N}^\infty \lvert a_n \rvert$ converges, and therefore $\sum_{n=1}^\infty \lvert a_n \rvert$ converges. Thus, $\sum_{n=1}^\infty a_n$ converges absolutely.

    \item $L > 1$:

      By the definition of $\limsup$, there exists a subsequence $(a_{n_k})$ such that
      \[%
        \sqrt[n_k]{\lvert a_{n_k} \rvert} \to L~\textrm{as}~k \to \infty
      .\]%
      Since $L > 1$, for sufficiently large $k$, we have
      \[%
        \sqrt[n_k]{\lvert a_{n_k} \rvert} > 1
      ,\]%
      which implies
      \[%
        \lvert a_{n_k} \rvert > 1^n = 1
      .\]%
      Thus, $\lvert a_{n_k} \rvert$ does not approach $0$, so the series $\sum_{n=1}^\infty a_n$ diverges.

    \item $L = 1$:

      The test is inconclusive because series with $L = 1$ may either converge or diverge. For instance
      \[%
        \sum_{n=1}^\infty \frac{1}{n^2}~\textrm{converges, while}~\sum_{n=1}^\infty \frac{1}{n}~\textrm{diverges}
      .\]%
      Both of these series satisfy $\limsup_{n \to \infty} \sqrt[n]{\lvert a_n \rvert} = 1$.\qedhere
  \end{enumerate}
\end{proof}

\begin{lemma}[Alternating Series Test]
  Let $(a_n)$ be a sequence satisfying the following requirements
  \begin{enumerate}
    \item $a_1 \ge a_2 \ge a_3 \ge \cdots \ge a_n \ge a_{n+1} \ge \cdots$,

    \item $a_n \to 0$,
  \end{enumerate}
  Then, the alternating series $\sum_{n=1}^\infty (-1)^{n+1} a_n$ converges.
\end{lemma}

\begin{proof}
  Let $S_n = \sum_{k=1}^n (-1)^{k+1} a_k$. Then, $S_{2n} = a_1 - a_2 + a_3 - a_4 + \cdots + a_{2n-1} - a_{2n}$. Since $a_n \to 0$, we have
  \[%
    \lvert S_{2n} - S_{2n-2} \rvert = \lvert a_{2n-1} - a_{2n} \rvert \to 0
  .\]%
  Thus, $(S_{2n})$ is a Cauchy sequence. Similarly, $(S_{2n+1})$ is also a Cauchy sequence. Thus, both sequences converge. Since the series is the sum of these two sequences, the series converges.
\end{proof}

\begin{lemma}[Condensation Test]
  Suppose $(b_n)$ is decreasing and satisfies $b_n \ge 0$ for all $n \in \N$. Then, the series $\sum_{n=1}^\infty b_n$ converges if and only if the series
  \[%
    \sum_{n=0}^\infty 2^nb_{2^n} = b_1 + 2b_2 + 4b_4 + 8b_8 + 16b_{16} + \cdots
  ,\]%
  converges.
\end{lemma}

\begin{proof}
  First, assume that $\sum_{n=0}^\infty 2^nb_{2^n}$ converges. Then, the partial sum
  \[%
    T_k = b_1 + 2b_2 + 4b_4 + \cdots + 2^kb_{2^k}
  ,\]%
  is bounded, by some number $M$. Because $b_n \ge 0$, we know that the partial sums are increasing. Let $S_m = b_1 + b_2 + b_3 + \cdots + b_m$. Fix $m$ and let $k$ be large enough to ensure $m \le 2_{k+1} - 1$. Then, $S_m \le S_{2^k + 1} - 1$ and
  \begin{align*}
    S_{2^{k+1}} - 1 &= b_1 + (b_2 + b_3) + (b_4 + \cdots + b_7) + \cdots + (b_{2^k} + \cdots + b_{2^{k+1} - 1}) \\
                    &\le b_1 + 2b_2 + 4b_4 + \cdots + 2^kb_{2^k} \\
                    &= T_k
  .\end{align*}
  Therefore, $S_m \le T_k \le M$ for all $m \in \N$. This implies that $(S_m)$ is bounded and therefore converges.
\end{proof}

This is a more general statement of the statement that the sum $\sum_{n=1}^\infty \frac{1}{n}$ diverges and that $\sum_{n=1}^\infty
\frac{1}{n^\rho}$ converges for $\rho > 1$.

\begin{corollary}[P-Series Test]
  The series
  \[%
    \sum_{n=1}^\infty \frac{1}{n^\rho}
  .\]%
  converges if and only if $\rho > 1$.
\end{corollary}

\begin{proof}
  Let $b_n = \frac{1}{n^\rho}$. By the Cauchy Condensation Test, the series $\sum_{n=1}^\infty b_n$ converges if and only if the series
  \[%
    \sum_{n=0}^\infty 2^n \cdot \frac{1}{(2^n)^\rho} = \sum_{n=0}^\infty \frac{1}{2^{n(\rho-1)}}
  .\]%
  converges. The latter is a geometric series with common ratio $r = \frac{1}{2^{\rho-1}}$. A geometric series converges if and only if $\lvert r \rvert < 1$. Thus, $\lvert r \rvert < 1$ implies
  \[%
    \frac{1}{2^{\rho-1}} < 1
  ,\]%
  which holds if and only if $\rho - 1 > 0$, or equivalently, $\rho > 1$. Therefore, $\sum_{n=1}^\infty \frac{1}{n^\rho}$ converges if and only if $\rho > 1$.
\end{proof}

The discrete analogue of integration by parts is often called \emph{summation by parts} or \emph{Abel's transformation}. It is the fundamental identity used to prove Dirichlet's test.

\begin{lemma}[Summation by parts]
  Let $(a_n)_{n \ge 1}$ and $(b_n)_{n \ge 1}$ be real sequences and set
  \[%
    B_n \coloneqq \sum_{k=1}^n b_k \quad (B_0 \coloneqq 0)
  .\]%
  For every integer $N\ge 1$ we have
  \[%
    \sum_{k=1}^N a_k b_k = a_N B_N + \sum_{k=1}^{N-1} (a_k - a_{k+1}) B_k
  .\]%
\end{lemma}

\begin{proof}
  Write each $b_k$ as a difference of partial sums, $b_k = B_k - B_{k-1}$. Then
  \[%
    \sum_{k=1}^N a_k b_k = \sum_{k=1}^N a_k (B_k - B_{k-1}) = \sum_{k=1}^N a_k B_k - \sum_{k=1}^N a_k B_{k-1}
  .\]%
  In the second sum shift the index $k \mapsto k + 1$ (noting $B_0 = 0$)
  \[%
    \sum_{k=1}^N a_k B_{k-1} = \sum_{k=0}^{N-1} a_{k+1} B_k = \sum_{k=1}^{N-1} a_{k+1} B_k
  .\]%
  Therefore
  \[%
    \sum_{k=1}^N a_k b_k = a_N B_N + \sum_{k=1}^{N-1} (a_k - a_{k+1}) B_k
  ,\]%
  as claimed.
\end{proof}

\begin{lemma}[Dirichlet's Test]
  Let $(a_n)$ and $(b_n)$ be sequences. Suppose
  \begin{enumerate}
    \item $(a_n)$ is monotone and $\lim_{n\to\infty} a_n = 0$,

    \item the partial sums $B_n=\sum_{k=1}^n b_k$ are bounded: there exists $M>0$ with $\abs{B_n} \le M$ for all $n$.
  \end{enumerate}
  Then the series $\sum_{n=1}^\infty a_n b_n$ converges.
\end{lemma}

\begin{proof}
  We give the standard proof using summation by parts. For clarity assume $(a_n)$ is nonincreasing (the other monotone cases are handled similarly by sign changes) and that $a_n\ge 0$ for all $n$; the general monotone-to-$0$ case is reduced to this by multiplying by $-1$ if necessary.

  Let $S_N \coloneqq \sum_{k=1}^N a_k b_k$ denote the $N$-th partial sum. By summation by parts,
  \[%
    S_N = a_N B_N + \sum_{k=1}^{N-1} (a_k - a_{k+1}) B_k
  .\]%
  By hypothesis there is $M>0$ with $\abs{B_k} \le M$ for every $k$. Because $(a_k)$ is nonincreasing and nonnegative, each difference $a_k-a_{k+1}\ge 0$, and the finite sum in the identity involves nonnegative weights. Estimate the absolute value of $S_N$ as follows:
  \[%
    \abs{S_N} \le \abs{a_N B_N} + \sum_{k=1}^{N-1} (a_k-a_{k+1}) \abs{B_k} \le M a_N + M\sum_{k=1}^{N-1} (a_k-a_{k+1})
  .\]%
  The telescoping sum equals
  \[%
    \sum_{k=1}^{N-1} (a_k-a_{k+1}) = a_1 - a_N
  .\]%
  Hence
  \[%
    \abs{S_N} \le M a_N + M(a_1 - a_N) = M a_1
  .\]%
  In particular the sequence $(S_N)$ of partial sums is bounded. To prove convergence we show it is Cauchy: for $N>n$ the same identity applied to $S_N-S_n$ gives
  \[%
    S_N - S_n = a_N B_N - a_n B_n + \sum_{k=n}^{N-1} (a_k-a_{k+1}) B_k
  ,\]%
  and therefore, using $\abs{B_k} \le M$,
  \[%
    \abs{S_N - S_n} \le M\abs{a_N} + M\abs{a_n} + M\sum_{k=n}^{N-1} (a_k-a_{k+1}) = M(\abs{a_N} + \abs{a_n} + a_n - a_N)
  .\]%
  Because $a_n \ge 0$ and $a_N \ge 0$ this simplifies to
  \[%
    \abs{S_N - S_n} \le 2M a_n
  .\]%
  Since $a_n \to 0$ as $n \to \infty$, for every $\epsilon > 0$ we can choose $n$ large enough that $2M a_n<\epsilon$, and then for all $N>n$ we have $\abs{S_N-S_n} < \epsilon$. Thus $(S_N)$ is Cauchy and hence convergent. This proves that $\sum_{n=1}^\infty a_n b_n$ converges.

  If $(a_n)$ is monotone but not nonnegative, write $a_n = a_n^+ - a_n^-$ with the monotone parts and apply the same argument to each part (or reduce to the nonnegative monotone case by multiplying by $-1$ when necessary). This completes the proof.
\end{proof}

\subsection{Double Summation of Infinite Series}

\begin{definition}[Iterated Summation]
  Given a doubly indexed array of real numbers $\{a_{ij} \mid i, j \in \N\}$, the \emph{iterated sum} of the array is defined as
  \[%
    S_{mn} = \sum_{i=1}^m \sum_{j=1}^n a_{ij}
  .\]%
  The \emph{double summation} is defined as
  \[%
    \sum_{i,j=1}^\infty a_{ij} = \lim_{m, n \to \infty} S_{mn}
  .\]%
\end{definition}

\begin{theorem}[Fubini's Theorem for Double Series]
  Let $\{a_{ij} \mid i, j \in \N\}$ be a doubly indexed array of real numbers.
  \begin{enumerate}
    \item If $\sum_{i,j=1}^\infty \lvert a_{ij} \rvert < \infty$ (absolute convergence), then the double series converges, and the order of summation does not matter
      \[%
        \sum_{i=1}^\infty \sum_{j=1}^\infty a_{ij} = \sum_{j=1}^\infty \sum_{i=1}^\infty a_{ij} = \sum_{i,j=1}^\infty a_{ij}
      .\]%

    \item If $\sum_{i=1}^\infty \sum_{j=1}^\infty a_{ij}$ converges conditionally (not absolutely), then rearranging the terms or changing the order of summation may alter the sum or even lead to divergence.
  \end{enumerate}
\end{theorem}

\begin{proof}
  Suppose $\sum_{i,j=1}^\infty \lvert a_{ij} \rvert < \infty$. Let $S_{mn} = \sum_{i=1}^m \sum_{j=1}^n a_{ij}$. For any fixed $m, n \in \N$, we can write
  \[%
    \lvert S_{mn} \rvert \leq \sum_{i=1}^m \sum_{j=1}^n \lvert a_{ij} \rvert \leq \sum_{i,j=1}^\infty \lvert a_{ij} \rvert
  .\]%
  By the monotone convergence theorem, the limits $\lim_{m \to \infty} \sum_{i=1}^m \sum_{j=1}^n a_{ij}$ and $\lim_{n \to \infty} \sum_{j=1}^n \sum_{i=1}^m a_{ij}$ exist and are equal. Thus,
  \[%
    \sum_{i=1}^\infty \sum_{j=1}^\infty a_{ij} = \sum_{j=1}^\infty \sum_{i=1}^\infty a_{ij}
  .\qedhere\]%
\end{proof}

\begin{corollary}[Order of Summation for Nonnegative Terms]
  If $a_{ij} \geq 0$ for all $i, j \in \N$, then
  \[%
    \sum_{i=1}^\infty \sum_{j=1}^\infty a_{ij} = \sum_{j=1}^\infty \sum_{i=1}^\infty a_{ij}
  ,\]%
  even if the series does not converge absolutely.
\end{corollary}

\begin{proof}
  For nonnegative terms, $\sum_{i,j=1}^\infty a_{ij}$ converges by monotonicity if and only if the partial sums $\sum_{i=1}^m \sum_{j=1}^n a_{ij}$ converge. Thus, the iterated sums also converge and are equal.
\end{proof}

\begin{example}[Harmonic Series Grid]
  Consider $a_{ij} = \frac{1}{i+j}$. We investigate the sums
  \[%
    \sum_{i=1}^\infty \sum_{j=1}^\infty a_{ij} \aand \sum_{j=1}^\infty \sum_{i=1}^\infty a_{ij}
  .\]%
  Since $a_{ij}$ is not absolutely convergent (due to divergence of $\sum_{i,j=1}^\infty \frac{1}{i+j}$), the series cannot be rearranged freely. Direct computation of either summation requires careful handling of the partial sums.
\end{example}

\begin{lemma}[Symmetry of Convergent Double Series]
  If $\{a_{ij}\}$ is absolutely convergent and symmetric, i.e., $a_{ij} = a_{ji}$ for all $i, j$, then
  \[%
    \sum_{i,j=1}^\infty a_{ij} = 2 \sum_{i=1}^\infty \sum_{j=i+1}^\infty a_{ij}
  .\]%
\end{lemma}

\begin{proof}
  Partition the terms into two symmetric halves: $a_{ij}$ for $i < j$ and $a_{ij}$ for $j < i$. Add the diagonal terms $a_{ii}$ separately. Absolute convergence ensures this rearrangement is valid.
\end{proof}

\subsection{Products of Infinite Series}

\begin{definition}[Cauchy Product of Infinite Series]
  Let $\sum_{n=0}^\infty a_n$ and $\sum_{n=0}^\infty b_n$ be two infinite series. The \emph{Cauchy product} of these series is defined as
  \[%
    \left(\sum_{n=0}^\infty a_n\right) \cdot \left(\sum_{n=0}^\infty b_n\right) = \sum_{n=0}^\infty \underbrace{\sum_{k=0}^n a_k b_{n-k}}_{c_n}
  .\]%
\end{definition}

\begin{theorem}[Cauchy Product Convergence]
  If $\sum_{n=0}^\infty a_n$ and $\sum_{n=0}^\infty b_n$ converge absolutely, then their Cauchy product also converges, and
  \[%
    \sum_{n=0}^\infty c_n = \left(\sum_{n=0}^\infty a_n\right) \cdot \left(\sum_{n=0}^\infty b_n\right)
  .\]%
\end{theorem}

\begin{proof}
  Since $\sum_{n=0}^\infty \lvert a_n \rvert < \infty$ and $\sum_{n=0}^\infty \lvert b_n \rvert < \infty$, the absolute convergence allows us to rearrange and group terms freely. Using the definition of the Cauchy product and the triangle inequality, we can write
  \[%
    \left\lvert\sum_{n=0}^\infty \sum_{k=0}^n a_k b_{n-k}\right\rvert \leq \sum_{n=0}^\infty \sum_{k=0}^n \lvert a_k \rvert \lvert b_{n-k} \rvert
  .\]%
  By Fubini's theorem for double series, the double sum can be rearranged into a product of two convergent series
  \[%
    \sum_{n=0}^\infty \sum_{k=0}^n \lvert a_k \rvert \lvert b_{n-k} \rvert = \left(\sum_{n=0}^\infty \lvert a_n \rvert\right) \cdot \left(\sum_{n=0}^\infty \lvert b_n \rvert\right)
  .\]%
  Thus, the Cauchy product converges absolutely, and the value is given by the product of the original sums.
\end{proof}

\begin{corollary}[Conditional Convergence of the Cauchy Product]
  If $\sum_{n=0}^\infty a_n$ and $\sum_{n=0}^\infty b_n$ converge conditionally (not absolutely), the Cauchy product may fail to converge or yield a different sum than the product of the original series.
\end{corollary}

\begin{example}[Harmonic Series and Cauchy Product]
  Let $a_n = \frac{1}{n+1}$ and $b_n = \frac{1}{n+1}$, corresponding to the harmonic series. Since the harmonic series diverges, their Cauchy product also diverges. However, truncating the series to $n$ terms and analyzing the partial sums provides useful approximations.
\end{example}

\begin{lemma}[Cauchy Product and Power Series]
  Let $\sum_{n=0}^\infty a_n x^n$ and $\sum_{n=0}^\infty b_n x^n$ be power series with radii of convergence $R_1$ and $R_2$, respectively. Then the Cauchy product of these series,
  \[%
    \sum_{n=0}^\infty c_n x^n, \quad c_n = \sum_{k=0}^n a_k b_{n-k}
  ,\]%
  converges for $\lvert x \rvert < \min(\{R_1, R_2\})$.
\end{lemma}

\begin{proof}
  Within the radius of convergence, the partial sums of the power series converge absolutely. Applying the Cauchy product formula term by term preserves convergence, as long as $\lvert x \rvert$ is strictly less than the minimum of the radii.
\end{proof}

\begin{example}[Product of Geometric Series]
  Consider the geometric series $\sum_{n=0}^\infty x^n = \frac{1}{1-x}$ for $\lvert x \rvert < 1$. The product of two such series is given by
  \[%
    \left(\sum_{n=0}^\infty x^n\right) \cdot \left(\sum_{n=0}^\infty x^n\right) = \sum_{n=0}^\infty c_n x^n
  ,\]%
  where
  \[%
    c_n = \sum_{k=0}^n 1 = n + 1
  .\]%
  Thus, the product is
  \[%
    \sum_{n=0}^\infty (n + 1)x^n = \frac{1}{(1 - x)^2}
  .\qedhere\]%
\end{example}

\begin{theorem}[Multiplication of Exponential Series]
  Let $e^x = \sum_{n=0}^\infty \frac{x^n}{n!}$. The product of two exponential series satisfies
  \[%
    \left(\sum_{n=0}^\infty \frac{x^n}{n!}\right) \cdot \left(\sum_{n=0}^\infty \frac{y^n}{n!}\right) = \sum_{n=0}^\infty \frac{(x+y)^n}{n!}
  .\]%
\end{theorem}

\begin{proof}
  Using the Cauchy product, the coefficient of $z^n$ in the product is given by
  \[%
    \sum_{k=0}^n \frac{x^k}{k!} \cdot \frac{y^{n-k}}{(n-k)!} = \frac{1}{n!} \sum_{k=0}^n \binom{n}{k} x^k y^{n-k}
  .\]%
  By the binomial theorem, this simplifies to
  \[%
    \frac{1}{n!} (x+y)^n
  .\]%
  Thus, the product of the exponential series is the exponential series of $x + y$.
\end{proof}
