\lecture{9}{Oct 18 2024 Fri (13:02:06)}{Monotone Convergence Theorem}

\begin{definition}[Monotone]
  A sequence $(a_n)$ is \emph{increasing} if $(\forall n \in \N)[a_n \le a_{n+1}]$ and \emph{decreasing} if $(\forall n \in \N)[a_n \ge a_{n+1}]$. A sequence is \emph{monotone} if it is either increasing or decreasing.
\end{definition}

\begin{theorem}[Monotone Convergence Theorem]
  Every bounded monotone sequence converges.
\end{theorem}

\begin{proof}
  Let $\left(a_n\right)$ be monotone and bounded. To prove $\left(a_n\right)$ converges using the definition of convergence, we are going to need a candidate for the limit. Let's assume the sequence is increasing (the decreasing case is handled similarly), and consider the set of points $\{a_n \mid n \in \N\}$. By assumption, this set is bounded, so we can let
  \[%
    s = \sup(\{a_n \mid n \in \N\})
  .\]%
  It seems reasonable to claim that $\lim a_n = s$.

  To prove this, let $\epsilon > 0$. Because $s$ is the least upper bound for $\{a_n \mid n \in \N\}$, $s - \epsilon$ is not an upper bound, so there exists a point in the sequence $a_N$ such that $s - \epsilon < a_N$. Now, the fact that $(a_n)$ is increasing implies that if $n \geq N$, then $a_N \leq a_n$. Hence,
  \[%
    s - \epsilon < a_N \leq a_n \leq s < s + \epsilon
  ,\]%
  which implies $\abs{a_n - s} <\epsilon$, as desired.
\end{proof}

We now have another tool to prove the Nested Interval Property.
\begin{theorem}[Nested Interval Property]
  For each $n \in \N$, assume we are given a closed interval $I_n = [a_n, b_n] = \left\{x \in \R \mid a_n \le x \le b_n\right\}$. Assume also that each $I_n$ contains the next interval, i.e., $I_n \supseteq I_{n+1}$. Then, the resulting sequence of closed intervals
  \[%
    I_1 \supseteq I_2 \supseteq I_3 \supseteq \cdots
  ,\]%
  has a nonempty intersection. That is,
  \[%
    \bigcap_{n=1}^\infty I_n \neq \emptyset
  .\]%
\end{theorem}

\begin{proof}
  Assume we have the following interval $I_n = [a_n, b_n]$ for each $n \in \N$. Since all $I_n \supseteq I_{n+1}$, then we have the following inequality
  \[%
    I_n = [a_n, b_n] \supseteq I_{n+1} = [a_{n+1}, b_{n+1}] \implies \begin{cases*}
      a_n \le a_{n+1} \\
      b_n \ge b_{n+1}
    \end{cases*}
  .\]%
  This means that $(a_n)$ is bounded by $b_1$ from above and $(b_n)$ is bounded by $a_1$ from below. Then, by the MCT, $(a_n)$ converges to $a$ and $(b_n)$ converges to $b$. Hence,
  \[%
    [a, b] \in \bigcap_{n=1}^\infty I_n
  .\qedhere\]%
\end{proof}

\begin{question}
  Let $S_1 = 1$ and $S_{n+1} = \sqrt{S_n + 1}$, for $n \ge 1$. Show $S_n$ converges and find the limit.
\end{question}

\begin{proof}[Solution]
  Let's find the first few terms of the sequence
  \[%
    S_1 = 1, \quad S_2 = \sqrt{2}, \quad S_3 = \sqrt{\sqrt{2} + 1}, \quad S_4 = \sqrt{\sqrt{\sqrt{2} + 1} + 1}
  .\]%
  Then, we get the following inequality
  \[%
    S_1 < S_2 < S_3 < S_4 < \cdots
  .\]%

  We can then use induction to show that $S_n < S_{n+1}$ for all $n \in \N$ and then show that the sequence is bounded. Then, we can conclude by the MTC that the sequence must converge.

  Using induction, assume $S_{n-1} \le S_n$. Now, we show that $S_n \le S_{n+1}$. We have
  \begin{align*}
    S_{n+1} - S_n &= \sqrt{S_n + 1} - \sqrt{S_{n-1} + 1} \\
                  &= \frac{(S_n + 1) - (S_{n-1} + 1)}{\sqrt{S_n + 1} + \sqrt{S_{n-1} + 1}} \\
                  &= \frac{S_n - S_{n-1}}{\sqrt{S_n + 1} + \sqrt{S_{n-1} + 1}} \ge 0
  .\end{align*}
  Therefore, $S_n \le S_{n+1}$ for all $n \in \N$, meaning that $S_n$ is increasing.

  We now show that $S_n$ is bounded. Assume $S_n \le 2$. Then $S_{n+1} = \sqrt{S_n + 1} \le \sqrt{2 + 1} = \sqrt{3} \le \sqrt{2}$.

  Hence, by induction, $S_n \le 2$ for all $n$. So, $(S_n)$ is an increasing bounded sequence. By MCT, $(S_n)$ converges to $L$.

  Now, all that's left is to find the limit of the sequence.
  \[%
    \lim_{n \to \infty} S_n = \lim_{n \to \infty} S_{n+1} = S \implies S = \sqrt{S + 1} \implies S^2 = S + 1
  .\]%
  Solving the equation using the quadratic equation, we get
  \[%
    S = \frac{1 \pm \sqrt{5}}{2}
  .\]%
  But, as we've seen before, the limit value of a sequence is always unique. So, how do we choose which one the limit is? We'll, it must be the positive term, since $S_n \ge 1$ for all $n \in \N$. Therefore, the limit of the sequence is
  \[%
    \lim_{n \to \infty} S_n = \frac{1 + \sqrt{5}}{2}
  .\qedhere\]%
\end{proof}
